---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# R.NURS

<!-- badges: start -->
<!-- badges: end -->

A Base R, simple implementation of the No-Underrun Sampler. This implementation aims to mostly directly implement the algorithm as described by the paper, with at most small changes for code aesthetics and performance. This version uses the memory saving technique described in section 2.2.

## Installation

You can install the development version of R.NURS from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("VisruthSK/R.NURS")
```

## Example: Neal's Funnel

```{r funnel}
library(R.NURS)
library(ggplot2)

set.seed(0)
logpdf_funnel <- function(theta) {
  y <- theta[1]
  dnorm(y, 0, 3, log = TRUE) + sum(dnorm(theta[-1], 0, exp(y / 2), log = TRUE))
}

samples <- NURS(
  logpdf_funnel,
  theta_init = rep(0, 15),
  n = 5000,
  epsilon = 0.1,
  h = 0.5,
  M = 5
)

data.frame(y = samples[, 1], x1 = samples[, 2]) |>
  ggplot(aes(x = y, y = x1)) +
  geom_point(alpha = 0.3) +
  theme_minimal()
```

## References

* Nawaf Bou-Rabee, Bob Carpenter, Sifan Liu, Stefan Oberdörster. 2025.
[The No-Underrun Sampler: A locally adaptive, gradient free, MCMC
method](https://arxiv.org/abs/2501.18548v2). *arXiv* 2501.18548 v2.

* Art B. Owen. 2013. [Monte Carlo theory, methods and examples](https://artowen.su.domains/mc/). *artowen.su.domains*.
